### 2016-07-04 related-work.tex
edits

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, by defining the
kinds of dependency cycles that are forbidden in each case. The
operational nature of Adya's specifications make them suitable for
runtime monitoring and anomaly
detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas the
declarative nature of our specifications make them suitable for formal
reasoning about program behaviour. Sivaramakrishnan \emph{et
  al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, and is thus not
suitable to reason operationally about a program as it builds the
trace; their framework also requires application invariants be
expressed in terms of the visibility relations used to define
isolation and consistency properties.  Cerone \emph{et
  al.}~\cite{gotsmanconcur15} present a framework to specify isolation
levels with atomic visibility, but their specifications are also for
complete traces; moreover, their formulation implicitly assumes
homogenity, \emph{i.e.}, they require all transactions to execute
using either {\sc si} or {\sc ser}, which is often not the case in
practice. While all the aforementioned specification frameworks use
the vocabulary introduced in~\cite{burckhardt14} to specify replicated
data types and different forms of weak consistency, none of them come
with an associated reasoning framework that can use such
specifications productively towards verifying programs under weak
isolation.

\paragraph{Reasoning under weak isolation} In~\cite{feketessi}, Fekete
\emph{et al.} propose a theory to characterize non-serializable
executions that arise under {\sc si}. Fekete~\cite{fekete2005} also
proposes an algorithm that allocates either {\sc si} or {\sc ser}
isolation levels to transactions while guaranteeing the
serializability of the execution. In~\cite{gotsmanpodc16}, Cerone
\emph{et al.} improve on Adya's {\sc si} specification and use it to
derive a static analysis that determines the safety of substituting
{\sc si} with a weaker variant called \iso{Parallel Snapshot
  Isolation}~\cite{psi}.  These efforts focus on establishing the
equivalence of executions between a pair of isolation levels, without
taking application invariants into account.  Bernstein \emph{et
  al.}~\cite{bern2000} propose informal semantic conditions to ensure
the satisfaction of application invariants under weaker isolation
levels.  However, no formal proof system is given.  Furthermore, these
techniques are tailor-made for a finite set of well-understood
isolation levels (rooted in~\cite{berenson}) under a pre-defined
store consistency model (usually {\sc sc}).

\paragraph{Reasoning under weak consistency} There have been several
recent proposals on reasoning techniques for programs executing under
weak consistency~\cite{bailisvldb, alvarocalm,
  gotsmanpopl16,redblueatc, redblueosdi, ecinec}. All of them assume a
system model that offers a choice between a \emph{coordination-free}
weak consistency level (\emph{e.g.}, {\sc ec}~\cite{redblueosdi,
  redblueatc, ecinec, alvarocalm, bailisvldb}) or causal
consistency~\cite{gotsmanpopl16}). The key technical issue in these
efforts involves proving that atomic and fully isolated operations
preserve application invariants when executed on these consistency
levels.  In contrast, we admit weakly isolated transactions, and our
system model accepts \emph{specifications} of consistency and
isolation levels drawn from an expressive logic.  In recent
work~\cite{gotsmanpopl16}, \iso{Parallel Snapshot Isolation} is
adapted to the aforementioned setting by interpreting it as a
consistency level that serializes writes to objects; a dedicated proof
rule is developed to help prove prove program invariants hold under
this model. By parameterizing our proof system over a gamut of weak
isolation specifications, we avoid the need to define a separate proof
rule for each new isolation level we may encounter.

\paragraph{Reasoning under relaxed memory} The reasoning mechanisms
used to describe and prove properties about weakly-isolated
transactions bear resemblance to those used to formalize relaxed
memory behaviour~\cite{battycpp}.  Ridge~\cite{rgtso} generalizes
rely-guarantee reasoning to the x86-TSO memory model.  Likewise,
Vafeiadis \emph{et al.}~\cite{rsl13} generalize concurrent separation
logic (CSL)~\cite{csl} to the C11 relaxed memory model.  Ferreira
\emph{et al.}~\cite{ferreira10} propose a parameterized operational
semantics for relaxed memory models, but the parameterization is over
a relation between relaxed memory programs and related {\sc sc}
programs. They also prove the soundness of CSL for the weakest memory
model expressible in their semantics, but neither a parameterized (in
the aforementioned sense) CSL, nor a CSL specialized for any relaxed
memory model is proposed.  ~\cite{DLZ+13} presents a \emph{buffered
  memory model} for Java that defines an axiomatic definition for the
JMM in terms of memory reorderings, and an operational instantiation
consistent with the TSO memory model found on x86 processors, and
demonstrates the equivalence of the two definitions.  However, there
is no attempt to parameterize the model over different reordering or
consistency definitions, which is a central focus of this paper.

>>>>

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, by defining the
kinds of dependency cycles that are forbidden in each case. The
operational nature of Adya's specifications make them suitable for
runtime monitoring and anomaly
detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas the
declarative nature of our specifications make them suitable for formal
reasoning about program behaviour. Sivaramakrishnan \emph{et
  al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, and is thus not
suitable to reason operationally about a program as it builds the
trace; their framework also requires application invariants be
expressed in terms of the visibility relations used to define
isolation and consistency properties.  Cerone \emph{et
  al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces, and
assumes all transactions execute using either {\sc si} or {\sc ser},
which is often not the case in practice. While all the aforementioned
specification frameworks use the vocabulary introduced
in~\cite{burckhardt14} to specify replicated data types, none of them
are equipped with a reasoning framework that can use such
specifications to verify programs under weak isolation.

\paragraph{Reasoning under weak isolation} In~\cite{feketessi}, Fekete
\emph{et al.} propose a theory to characterize non-serializable
executions that arise under {\sc si}. Fekete~\cite{fekete2005} also
proposes an algorithm that allocates either {\sc si} or {\sc ser}
isolation levels to transactions while guaranteeing the
serializability. In~\cite{gotsmanpodc16}, Cerone \emph{et al.} improve
on Adya's {\sc si} specification and use it to derive a static
analysis that determines the safety of substituting {\sc si} with a
weaker variant called \iso{Parallel Snapshot Isolation}~\cite{psi}.
These efforts focus on establishing the equivalence of executions
between a pair of isolation levels, without taking application
invariants into account.  Bernstein \emph{et al.}~\cite{bern2000}
propose informal semantic conditions to ensure the satisfaction of
application invariants under weaker isolation levels.  All these
techniques are tailor-made for a finite set of well-understood
isolation levels (rooted in~\cite{berenson}) under a pre-defined store
consistency model (usually {\sc sc}).

\paragraph{Reasoning under weak consistency} There have been several
recent proposals to reason about programs executing under weak
consistency~\cite{bailisvldb, alvarocalm, gotsmanpopl16,redblueatc,
  redblueosdi, ecinec}. All of them assume a system model that offers
a choice between a \emph{coordination-free} weak consistency level
(\emph{e.g.}, eventual consistency~\cite{redblueosdi, redblueatc,
  ecinec, alvarocalm, bailisvldb}) or causal
consistency~\cite{LBC16,gotsmanpopl16}). The key technical issue in
these efforts involves proving that atomic and fully isolated
operations preserve application invariants when executed on these
consistency levels.  In contrast, we admit weakly isolated
transactions, and our system model accepts \emph{specifications} of
consistency and isolation levels drawn from an expressive logic.
Gotsman \emph{et al.}~\cite{gotsmanpopl16} adapts \iso{Parallel
  Snapshot Isolation} to the aforementioned setting by interpreting it
as a consistency level that serializes writes to objects; a dedicated
proof rule is developed to help prove prove program invariants hold
under this model. By parameterizing our proof system over a gamut of
weak isolation specifications, we avoid the need to define a separate
proof rule for each new isolation level we may encounter.

\paragraph{Reasoning under relaxed memory} The reasoning mechanisms
used to describe and prove properties about weakly-isolated
transactions bear resemblance to those used to formalize relaxed
memory behaviour~\cite{battycpp}.  Ridge~\cite{rgtso} generalizes
rely-guarantee reasoning to the x86-TSO memory model.  Likewise,
Vafeiadis \emph{et al.}~\cite{rsl13} generalize concurrent separation
logic (CSL)~\cite{csl} to the C11 relaxed memory model.  Ferreira
\emph{et al.}~\cite{ferreira10} propose a parameterized operational
semantics for relaxed memory models, but the parameterization is over
a relation between relaxed memory programs and related {\sc sc}
programs. Demange \emph{et al.}~\cite{DLZ+13} present a \emph{buffered
  memory model} for Java that defines an axiomatic definition for the
JMM in terms of memory reorderings, and an operational instantiation
consistent with the TSO memory model, and demonstrate the equivalence
of the two definitions.
